---
title: "Some Examples on R Notebook"
author: "Akihito Kamata"
date: "January 17, 2020"
output: html_notebook
---

### Preparation
Load required pachages.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
library(psych)
library(sjmisc)
library(sjPlot)
```

Read SPSS data file. `haven` package is required for `read_spss` function.
```{r}
dat10 <- read_spss("Data/Chapman.sav")
```

### Descriptive Statistics
Run descriptives for selected variables: Use `describe` in `psych` package.
```{r}
dat10 %>% select(SYSBP, DIABP, CHOLES) %>% describe(fast=T)
```

The mean and the standard deviation of systolic blood pressure was $M=121.64$ and $sd=16.70$. 

Run descriptives by groups: Use `describeBy` in `psych` package.
```{r}
dat10 %>% select(SYSBP, DIABP, CHOLES) %>% describeBy(group=dat10$CORON, fast=T)
```

Run frequency table for categorical variables by "frq" function in the `sjmisc` package. 
```{r}
dat10 %>% select(CORON) %>% frq()
```

### Plots
Scatter plot with fitted line.
```{r}
dat10 %>% 
  ggplot(aes(x=SYSBP, y=DIABP)) + 
  geom_point() +
  theme_bw() +
  stat_smooth(method = "lm", formula = y ~ x)
```

Bargraphs for Likert-type scale items.
```{r}
lik.dat <- read_csv("Data/likert.csv")
labels01 <- c("Very dissatisfied",
              "Somewhat dissatisfied",
              "Neither satisfied or dissatisfied",
              "Somewhat satisfied",
              "Very satisfied")
lik.dat %>% select(item1:item5) %>%  
  plot_likert(cat.neutral = 3,
              grid.range = c(1.5, .5),
              values = "sum.outside",
              legend.labels = labels01,
              show.prc.sign = T) + theme_bw()
```

